Introductory statistics with `intRo`

### 2016-06-06 jcgs-paper-2016/response.Rmd

---
title: "Response to reviewers"
author: "Eric Hare, Andee Kaplan"
date: "May 26, 2016"
output: html_document
---

# Reviewer 1

(1) The paper seems to me inappropriate for the JCGS. Because it focuses purely on software, I think that (once the other issues are addressed) it would be more appropriate for the Journal of Statistical Software or the R Journal.

**We appreciate the reviewer's perspective on the context in which our paper fits. While we are focusing on a software system, the aim of JCGS includes to "extend the use of computational and graphical methods in statistics and data analysis". It also contains "more on numerical graphical displays and methods, and perceptions." While this is an educational piece of software, this paper focuses on more general computational concepts of reproducibility, reactivity, and modularity, in the context of bringing statistical content to an end user. As such, we feel it does have a place within JCGS.**

@@

(2) Although the software looks promising, unless I've missed something, it currently has very limited functionality -- indeed, not enough to sustain most introductory statistics courses. This isn't a permanent limitation, however, since the authors have created interesting infrastructure for extending the software. In its current state, the software doesn't quite look ready for prime time.

**Our intent was not for `intRo` to sustain an introductory statistics course, but rather to supplement it. In other words, limited functionality was actually a design goal from the beginning so as not to overwhelm a student seeing statistics for the first time. While we agree with the reviewer that it is likely there will be course topics that `intRo` doesn't support, if additional functionality is desired by the instructor, more modules can be added to support this. We have extended the discussion of modularity in intRo to further emphasize this idea.**

@@

(3) The authors ignore other (and at present, much more capable) GUIs for R -- the R Commander and Deducer. Since these are the most direct competitors to intRo -- more so than, e.g., Excel and JMP -- it seems obvious that the authors should address similarities, differences, advantages, and disadvantages. One obvious difference is that these other GUIs run on individual computers rather than being web-based, and it would also make sense to address more generally the advantages and disadvantages of the two approaches. (For example, in my tests of intRo, I noticed slow response, even with very small data sets. Of course, it's hard to know to what to attribute that -- it could as easily have been a network issue as slowness at the server end.)

**We have included references and a brief discussion on some other GUIs for R, including R Commander and Deducer. We have described some of the differences in our approach to providing a front-end. As for the slow-response time in intRo, this can certainly occur in a couple of different scenarios. One would be if the student is working with a large dataset, as mentioned in Section 5. A second reason for slowdown would be if the student's internet connection is performing slowly. This is an unfortunate limitation of web-based interactive software. However, we have detailed the process by which a student may download and run an intRo instance locally to counteract this possibility.**

@@

I think the authors need to decide, “What sort of paper am I writing?” Included in the manuscript are elements of “a paper about teaching introductory statistics using a software system”, elements of “a paper about designing and building a simple extensible system”, and “a paper about how to use and extend a particular system”. I think the most compelling story for the Journal of Computational and Graphical Statistics is the designing and building of a simple online system for intro stat that is easy to learn, use and extend – and teaches R as a side effect. This would leave room for a sister publication in something like the Journal of Statistics Education which actually focusses on teaching aspects. If you buy this conception, the teaching-related aspects should be shortened, shaved of much detail and form the basis of an introductory motivational section that leads into a design brief and a plan of attack (using shiny etc). There needs to be high-level discussion of how the system is conceived and put together accompanied by high-level diagrams before descending into code fragments (much detailed code-fragment discussion could go to appendices). The paper just needs to convey the sense of how things are done. Actual details of how to customise and extend the system are provided in the system’s documentation.

**We have taken care to significantly shift the focus away from the teaching aspects, and towards the extensible software system aspects. In particular, we have removed the lengthy teaching discussion and moved elements of it to the motivational section. We have also trimmed much of the content involving supported and unsupported functionality. Specific changes are noted in the responses to the further points raised below.**

Everywhere – distinguish between user=student (who uses what is served up) and user=teacher (who customises what is served up)

**We went through and replaced instances of "user" with either "student" or "teacher" as appropriate.**

@@

p. 2: Another online analysis system implemented in shiny is iNZight Lite at http://lite.docker.stat.auckland.ac.nz/. It is more ambitious and therefore more complex. Also it does not have the “exposing the R code” capability.

**We have included a mention of iNZight Lite, along with the shared goals and some differences with intRo.**

@@

R Commander should enter the discussion around here (even though it isn’t online)

**We have included R Commander in the discussion.**

P 5: Fig 2 shows a simple design that should be easy for casual users (as intro students really are) to get their heads around. Not all that keen on the use of “statistics” as a category heading in Panel 2 because it reinforces the ghettoizing of the word conflicting with a message that says everything you are doing here is statistics.

**We have updated the application (and the screenshot) to refer to the contingency, regression, and t-test modules category heading as "inference" rather than "statistics", to help emphasize that numerical and graphical summaries are also statistics.**

@@

P 8: I think this descends to details too quickly. The big features are designing for extensibility, with conceptions about switching modules on and off and adding modules, and the interface as presented being assembled on the fly when the system starts up.

**We agree. We have restructured this portion of the paper to merely discuss the motivation behind an extensible system, with mentions of disabling and enabling modules. We have moved the details to the Understanding intRo section with the details on modules.**

P 9-10 looks like appendix stuff to me.

**We have move the implementation details of dynamic UI generation to the appendix.**

@@

P 14 Last para Section 3: could go into motivational discussion.

**We have moved this paragraph to the motivational discussion.**

Section 4.1 is a big part of what I see as the real story but descends to details too quickly.

**We have eased the transition into this section (now Section 3.1) by talking more generally about the benefits behind a modular structure. We also have made sure to clarify that our concept of modules is not to be confused with Shiny's new modules feature**

### 2016-06-06 jcgs-paper-2016/paper.Rmd
Update response and paper based on meeting

Three fundamental philosophies that guided the creation of `intRo`. In particular, `intRo` is \emph{easy} to use and can be an \emph{exciting} part of learning statistics. Additionally, `intRo` is an \emph{extensible} tool, allowing for a user of `intRo` to tailor the tool for his or her own classroom needs.

In the development of `intRo`, we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training. We used large, easy to click icons in the page header to help students find what they need more easily. We also made the functionality available the minimal necessary for an introductory statistics course. Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in `intRo`. In this instance, a user clicks on the Graphical tab to create a mosaic plot. The user sees the plot, and elects to click the save button to store the plot and its corresponding code to the final compendium.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/user_experience.pdf}
\caption{A typical user experience of generating a result in `intRo`. In this instance, a user clicks on the Graphical tab to create a mosaic plot. The user sees the plot, and elects to click the save button to store the plot (and its corresponding code) to the final compendium.}
\label{fig:user_experience}
\end{figure}

Beyond being simple, `intRo` is also consistent. The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules. To the user, a module is simply a page of statistics functionality that maintains a consistent layout, helping the user to become familiar with the location of the options, the results, and the code. Figure \ref{fig:ui} highlights the five elements that comprise the `intRo` interface.

\begin{figure*}[ht!]
\centering
\includegraphics[width=\linewidth]{images/ui_annotate.pdf}
\caption{The five elements that comprise the `intRo` application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
\label{fig:ui}
\end{figure*}
 1. **Top Navigation** - The top navigation bar includes two sets of clickable icons. The left-aligned buttons are informational buttons. The first is a link to `intRo`. The second is a link to the documentation page. The third is a link to the GitHub repository where the code for `intRo` is housed. The final button is a link to our websites, which contain contact information if there are any questions or comments. The right-aligned buttons are `intRo` utilities. The first is a link to toggle the visibility of the code panel (5). The middle icon downloads an rmarkdown [@rmarkdown] document of the analysis performed. The last is a link to print the stored module results, and the associated code (if visible).
 2. **Side Navigation** - The side navigation panel includes a list of data analysis tasks.
 3. **Options Panel** - The options panel includes task-specific options which the user can use to customize their results.
 4. **Results Panel** - The results pane displays the result of the selected module and options.
 5. **Code Panel** - The code panel displays the R code used to generate the results from the student's `intRo` session. The code panel is shown by default to facilitate a transition to coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.

@@

Statistics courses with a lab component can most benefit from the use of `intRo`. As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries. They simply upload a dataset and click the module appropriate to the section of the class. Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in. Early labs can walk the user through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the user can easily navigate the `intRo` interface to obtain the results.

`intRo` has an ulterior motive as well: to get students excited about programming. By navigating about the user interface of `intRo`, students are actually creating a fully-executable R script that they can download and run locally as well as viewing the script change real-time within the application. This code creation element of `intRo` is meant to generate excitement about programming in R and empower students to feel that they can generate code as well. `intRo` uses rmarkdown's render function in order to print the results, by dymanically executing the user's R script. By default, the output will include the R code, but if the user elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results. This reproducibility framework has a number of key advantages:

1. This eases a user who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code. Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many inexperienced users may have.

@@

An `intRo` module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality. `intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework. In this section, we detail the structure of these modules, and the process of creating and deploying new modules.

\begin{figure*}[ht!]
\centering
\includegraphics[width=\linewidth]{images/app_creation_modules.pdf}
\caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within `intRo`.}
\label{fig:app_creation_modules}
\end{figure*}

`intRo`'s user interface and functionality is dynamically generated depending on the set of modules enabled. This process is detailed in figure \ref{fig:app_creation_modules}. The key driver to populating \texttt{server.R} and \texttt{ui.R} is the navigation file \texttt{modules.txt}, which defines the placement of each module. The interface is then created with the following statement.

\footnotesize
```{r ui, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
## Source ui
module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")
sapply(file.path("modules", dir("modules")[dir("modules") != "modules.txt"], "ui.R"), source)

\footnotesize
```{r server, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
shinyServer(function(input, output, session) {
  ## Module info
  module_info <- read.table("modules/modules.txt", header = TRUE, sep=",")

  ## Modules
  types <- c("helper.R", "static.R", "observe.R", "reactive.R", "output.R")
  modules_tosource <- file.path("modules", 
                                apply(expand.grid(module_info$module, types), 1, paste, collapse = "/"))

  ## Source the modules
  for (mod in modules_tosource) {
      source(mod, local = TRUE)
  }
})
```
\normalsize
In this way, we were able to have `intRo` be fully extensible, its structure and functionality dependent entirely on the modules present within the application.

A module consists of the following scripts:

* *helper.R* - R code that performs some statistical analysis or transformation. This would typically be in the form of a function, and similar to any standard R script.
* *libraries.R* - Code to load any libraries which are not part of core R.
* *observe.R* - Shiny observer code typically used to update choices of an input box.
* *output.R* - Shiny output code defining the results of the analysis that should be displayed to the user.
* *reactive.R* - Shiny reactives, typically containing data that depend on inputs.
* *ui.R* - Shiny user interface definition, including the placement of the inputs and outputs.

@@

This script defines all the inputs and outputs that the user will see. The only requirements from `intRo`'s perspective are (1) that there exist a store button at the bottom of the middle panel for storing the results of the analysis in the code script, and (2) that configuration options appear in the width 4 column in the middle, and output appears in the width 8 column on the right. The remaining input and output definitions depend on the statistical analysis or transformation being performed.

@@

intRo leverages the reactive programming nature of Shiny, and as such is designed around the idea of user input cascading through the entire application. In a typical Shiny application, users interact with inputs that act as parameters to function, which in turn yield different results. Within `intRo`, the users are able to interact with and manipulate the data underlying the entire application. This posed many challenges in the creation of `intRo` and drove design decisions, namely timely save points according to the user's workflow, and reactive updating of variable lists tied to inputs across the entire application. Because the user may experiment with different configurations or select different variables, we did not want to store all actions taken in the intRo session. Rather, each module includes a button allowing the user to explicitly store the output visible in the results panel into the R script. This way, output is only stored when the user is satisfied, and the resulting output is not cluttered with unnecessary information.

In the creation of `intRo` we walked a fine line between giving the user flexibility and having realistic usability. At the same time, `intRo` was created as a consumer of another package, Shiny, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability. For a tangible example, consider the graphical summaries module. We only allow variables of a type consistent with the selected plot to be displayed. This is a conscious decision that limits an `intRo` user's flexibility, while maximizing the usability (by minimizing crashes) of the application. On the flip side of this, Shiny allows much higher flexibility. For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within `intRo`. However, Shiny does have limits on its flexibility based on the designers decisions for usability. One current example is the slider element. This element allows for fixed width steps from its minimum to its maximum. The JavaScript library being utilized in Shiny allows for arbitrary function calls to to generate these steps, however they must be written in plain JavaScript. This is an example of a decision made by the developers of Shiny to limit functionality in favor of usability of their package.

@@

Although users can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to download, customize, and deploy their own instance, perhaps with new modules or modified theming or functionality. `intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`. Currently, the package is only available on GitHub, and can be installed using the devtools package as follows:

@@

If the intent is to use a specific instance of `intRo` where many users will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL. The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package. Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the user's ShinyApps.io account. The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme. It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID. It also takes `...` as additional arguments to be passed into the `deployApp` routine. For example, if we wished to deploy the instance of `intRo` we ran previously, we would call it like so:

@@

In addition to the current functionality, there are some practical improvements in the works that will make `intRo` more useful to both students and instructors. In particular, we have begun development on an R package which will allow `intRo` modules to be created automatically from user written R code. This package will generate the necessary file structure to allow the module's incorporation into `intRo` as well as translate user code to `intRo` compatible code and populate the necessary files. This will vastly improve `intRo`'s flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses. Additionally, we would like to utilize the expanded interactive capabilities of ggvis in order to make \texttt{intRo's} plots more engaging to students. One way to do this would be implementing linked plots, in which interactions with one plot are reflected in other plots that illustrate the same data. This would be particularly useful in the regression module so that students could explore observations with high influence and leverage.

@@

>>>>

Three fundamental philosophies that guided the creation of `intRo`. In particular, `intRo` is \emph{easy} to use and can be an \emph{exciting} part of learning statistics. Additionally, `intRo` is an \emph{extensible} tool, allowing for a course instructor using `intRo` to tailor the tool for his or her own classroom needs.

In the development of `intRo`, we focused on aspects of the user interface (UI) and output that make it easy to pick up without extensive training. We used large, easy to click icons in the page header to help students find what they need more easily. We also made the functionality available the minimal necessary for an introductory statistics course. Figure \ref{fig:user_experience} illustrates the simple steps a student takes to generate a result in `intRo`. In this instance, a student clicks on the Graphical tab to create a mosaic plot. The student sees the plot, and elects to click the save button to store the plot and its corresponding code to the final compendium.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{images/user_experience.pdf}
\caption{A typical student experience of generating a result in \texttt{intRo}. In this instance, a student clicks on the Graphical tab to create a mosaic plot. The student sees the plot, and elects to click the save button to store the plot (and its corresponding code) to the final compendium.}
\label{fig:user_experience}
\end{figure}

Beyond being simple, `intRo` is also consistent. The tool is organized around specific tasks a student may perform in the process of a data analysis, called modules. To the student, a module is simply a page of statistics functionality that maintains a consistent layout, helping the student to become familiar with the location of the options, the results, and the code. Figure \ref{fig:ui} highlights the five elements that comprise the `intRo` interface.

\begin{figure*}[ht!]
\centering
\includegraphics[width=\linewidth]{images/ui_annotate.pdf}
\caption{The five elements that comprise the \texttt{intRo} application: 1) top navigation, 2) side navigation, 3) options panel, 4) results panel, and 5) code panel.}
\label{fig:ui}
\end{figure*}
 1. **Top Navigation** - The top navigation bar includes two sets of clickable icons. The left-aligned buttons are informational buttons. The first is a link to `intRo`. The second is a link to the documentation page. The third is a link to the GitHub repository where the code for `intRo` is housed. The final button is a link to our websites, which contain contact information if there are any questions or comments. The right-aligned buttons are `intRo` utilities. The first is a link to toggle the visibility of the code panel (5). The middle icon downloads an rmarkdown [@rmarkdown] document of the analysis performed. The last is a link to print the stored module results, and the associated code (if visible).
 2. **Side Navigation** - The side navigation panel includes a list of data analysis tasks.
 3. **Options Panel** - The options panel includes task-specific options which the student can use to customize their results.
 4. **Results Panel** - The results pane displays the result of the selected module and options.
 5. **Code Panel** - The code panel displays the R code used to generate the results from the student's `intRo` session. The code panel is shown by default to facilitate a transition to coding, but can be hidden by clicking the code toggle button in the Top Navigation bar.

@@

Statistics courses with a lab component can most benefit from the use of `intRo`. As students learn the aforementioned concepts, they can practice using the tools to perform the statistical tests and summaries. They simply upload a dataset and click the module appropriate to the section of the class. Once they choose the options desired by the instructor (variables, confidence levels, etc.), they can view the results and ultimately print the output for turning in. Early labs can walk the student through the limited number of button presses and clicks necessary, while later labs can more vaguely describe the intended output in hopes that the student can easily navigate the `intRo` interface to obtain the results.

`intRo` has an ulterior motive as well: to get students excited about programming. By navigating about the user interface of `intRo`, students are actually creating a fully-executable R script that they can download and run locally as well as viewing the script change real-time within the application. This code creation element of `intRo` is meant to generate excitement about programming in R and empower students to feel that they can generate code as well. `intRo` uses rmarkdown's render function in order to print the results, by dymanically executing the student's R script. By default, the output will include the R code, but if the student elects to hide the source code by clicking the code toggle button at the top, the code will not appear in the printed results. This reproducibility framework has a number of key advantages:

1. This eases a student who may be intimidated by programming into the idea that interacting with a user interface is really just a frontend for code. Seeing the correspondence between graphical clicks and printed code will hopefully also lessen the fear of coding that many students may have.

@@

An `intRo` module is a set of self-contained executible R scripts that together produce a set of introductory statistics functionality. `intRo` modules were designed in this way to allow for simple dynamic creation of the user interface at run-time, as well as ease the process of converting existing analysis code to the `intRo` framework. A high-level diagram of this process is given in figure \ref{fig:app_creation_modules}. Based on the modules contained within the `modules` folder, the interface and functionality of the app is dynamically generated at runtime.

\begin{figure*}[ht!]
\centering
\includegraphics[width=\linewidth]{images/app_creation_modules.pdf}
\caption{This figure depicts how the \texttt{server.R} and \texttt{ui.R} files are populated using the modular structure within \texttt{intRo}.}
\label{fig:app_creation_modules}
\end{figure*}

Section \ref{dynamic-ui-generation} provides some technical details on how we implemented this. For the rest of this section, we focus on the structure and development of the modules themselves, to aid in the process of creating and deploying new modules.

Modularity was a design decision we focused on from the start of `intRo's` development. There are some practical benefits to thinking of related statistics and data science functionality in terms of modules:

1. Course instructors can easily enable or disable functionality depending on the exact needs of the particular class being taught. Because this is done at run-time, including new modules is as simple as downloading and placing it within 'intRo's` modules folder, or removing existing modules from that folder.
2. Errors can be more easily isolated to specific components. For instance, if an error is encountered, simply disabling the module can provide a temporary workaround while the issue is identified.
3. Modularity helps to organize the different pieces of code into functionality chunks that make it easier for developers to maintain.

`intRo` modules are not to be confused with Shiny modules [@shinymodules]. Shiny modules are a recent feature added to Shiny which allows the bundling of inputs and outputs into a single set of functionality. They are more general and suitable for any application. `intRo` modules are specifically for statistics functionality and work within the `intRo` application only.

An `intRo` module consists of the following scripts:

* *helper.R* - R code that performs some statistical analysis or transformation. This would typically be in the form of a function, and similar to any standard R script.
* *libraries.R* - Code to load any libraries which are not part of core R.
* *observe.R* - Shiny observer code typically used to update choices of an input box.
* *output.R* - Shiny output code defining the results of the analysis that should be displayed to the student.
* *reactive.R* - Shiny reactives, typically containing data that depend on inputs.
* *ui.R* - Shiny user interface definition, including the placement of the inputs and outputs.

@@

This script defines all the inputs and outputs that the student will see. The only requirements from `intRo's` perspective are (1) that there exist a store button at the bottom of the middle panel for storing the results of the analysis in the code script, and (2) that configuration options appear in the width 4 column in the middle, and output appears in the width 8 column on the right. The remaining input and output definitions depend on the statistical analysis or transformation being performed.

@@

intRo leverages the reactive programming nature of Shiny, and as such is designed around the idea of user input cascading through the entire application. In a typical Shiny application, users interact with inputs that act as parameters to function, which in turn yield different results. Within `intRo`, the students are able to interact with and manipulate the data underlying the entire application. This posed many challenges in the creation of `intRo` and drove design decisions, namely timely save points according to the student's workflow, and reactive updating of variable lists tied to inputs across the entire application. Because the student may experiment with different configurations or select different variables, we did not want to store all actions taken in the intRo session. Rather, each module includes a button allowing the student to explicitly store the output visible in the results panel into the R script. This way, output is only stored when the student is satisfied, and the resulting output is not cluttered with unnecessary information.

In the creation of `intRo` we walked a fine line between giving the student flexibility and having realistic usability. At the same time, `intRo` was created as a consumer of another package, Shiny, in which we as developers were the beneficiaries of another team of developers' decision to balance flexibility and usability. For a tangible example, consider the graphical summaries module. We only allow variables of a type consistent with the selected plot to be displayed. This is a conscious decision that limits an `intRo` user's flexibility, while maximizing the usability (by minimizing crashes) of the application. On the flip side of this, Shiny allows much higher flexibility. For instance, the entire application (including user interface) is created dynamically upon load, based on the modules currently housed within `intRo`. However, Shiny does have limits on its flexibility based on the designers decisions for usability. One current example is the slider element. This element allows for fixed width steps from its minimum to its maximum. The JavaScript library being utilized in Shiny allows for arbitrary function calls to to generate these steps, however they must be written in plain JavaScript. This is an example of a decision made by the developers of Shiny to limit functionality in favor of usability of their package.

@@

Although students can access intRo from \url{http://www.intro-stats.com}, course instructors may wish to download, customize, and deploy their own instance, perhaps with new modules or modified theming or functionality. `intRo` can be downloaded, ran, and deployed on ShinyApps.io through the use of the R package `intRo`. Currently, the package is only available on GitHub, and can be installed using the devtools package as follows:

@@

If the intent is to use a specific instance of `intRo` where many students will access it at the same time, such as in an introductory statistics class, it may be preferable to deploy a custom instance of `intRo` to a publicly accessible URL. The package provides a function `deploy_intRo` which is a wrapper for the `deployApp` function contained in the shinyapps package. Once the shinyapps package is installed and configured, `deploy_intRo` will upload `intRo` as an application on the instructor's ShinyApps.io account. The function takes the same arguments as `run_intRo`, so it can be deployed with a custom selection of modules, and a customized theme. It also takes an additional argument `google_analytics`, which allows the specification of a Google Analytics tracking ID. It also takes `...` as additional arguments to be passed into the `deployApp` routine. For example, if we wished to deploy the instance of `intRo` we ran previously, we would call it like so:

@@

In addition to the current functionality, there are some practical improvements in the works that will make `intRo` more useful to both students and instructors. In particular, we have begun development on an R package which will allow `intRo` modules to be created automatically from user written R code. This package will generate the necessary file structure to allow the module's incorporation into `intRo` as well as translate user code to `intRo` compatible code and populate the necessary files. This will vastly improve `intRo's` flexibility and allow it to be used in a wider range of curricula, including more advanced statistics courses. Additionally, we would like to utilize the expanded interactive capabilities of ggvis in order to make \texttt{intRo's} plots more engaging to students. One way to do this would be implementing linked plots, in which interactions with one plot are reflected in other plots that illustrate the same data. This would be particularly useful in the regression module so that students could explore observations with high influence and leverage.

@@

# Appendix

## Dynamic UI Generation

`intRo's` user interface and functionality is dynamically generated depending on the set of modules enabled. The key driver to populating \texttt{server.R} and \texttt{ui.R} is the modules folder, the directory structure of which defines the placement of each module. The interface is then created with the following statement.

\footnotesize
```{r ui, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
## Source ui
mylist <- list()
old_heading <- ""
for (i in seq_along(modules)) {
    my.module <- strsplit(modules[i], "/")[[1]]
    if (my.module[1] != old_heading) {
        mylist[[length(mylist) + 1]] <- Hmisc::capitalize(my.module[1])
        old_heading <- my.module[1]
    }
    mylist[[length(mylist) + 1]] <- get(paste(my.module[2], "ui", sep = "_"))
}

\footnotesize
```{r server, eval=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=25)}
shinyServer(function(input, output, session) {
    types <- c("helper.R", "observe.R", "reactive.R", "output.R")

    modules_tosource <- file.path("modules", apply(expand.grid(modules, types), 1, paste, collapse = "/"))

    for (mod in modules_tosource) {
        source(mod, local = TRUE)
    }
})
```
\normalsize
In this way, we were able to have `intRo` be fully extensible, its structure and functionality dependent entirely on the modules present within the application.
